Air Quality and CMS Data

### 2021-01-25 code/data-analysis/factor-analysis.Rmd
1 to 5yr mvavg modify smooth function

#using smoothed.yr function
smooth.out <- mapply(Smoothed.Year, site.para$Site.ID, site.para$Parameter.Symbol)
para.monthly <- data.frame(Site.ID=unlist(smooth.out[1,]),
                           Parameter.Symbol=unlist(smooth.out[2,]),
                           Date=unlist(smooth.out[3,]), 
                           para.1yr=unlist(smooth.out[4,]))

>>>>

#using smoothed.yr function
smooth.out <- mapply(Smoothed.Year, 
                     site.para$Site.ID, site.para$Parameter.Symbol, 0.99, 350)
para.monthly <- data.frame(Site.ID=unlist(smooth.out[1,]),
                           Parameter.Symbol=unlist(smooth.out[2,]),
                           Date=unlist(smooth.out[3,]), 
                           para.1yr=unlist(smooth.out[4,]))

### 2021-01-25 code/data-processors/air-data-processor.Rmd
1 to 5yr mvavg modify smooth function

# pmc-site-zip-bz12.csv was uploaded on NEU Cluster to be merged with CMS data
# It has 320 sites with 8446 zipcodes around them
```

@@

>>>>

# pmc-site-zip-bz12.csv was uploaded on NEU Cluster to be merged with CMS data
# It has 320 sites with 8446 zipcodes around them
# 317 out of 320 sites have population living in their 12-mile bufferzone

# pmc-site-zip-bz6.csv was uploaded on NEU Cluster to be merged with CMS data
# 271 out of 274 sites have population living in their 6-mile bufferzone
# We extended the bufferzone around these 271 sites to 12- and 24- miles too
```

@@

```{r Smoothing Data & Caluculating 1 to 5-yr mvavg for each month}
cmp.sub <- c("AS", "CA", "CL", "NA", "NI", "NO3", "OC", 
             "PB", "PM", "SI", "SO4", "V", "ZN")
dt.air <- dt.pmc.daily %>% 
  select(Site.ID, Date, cmp.sub) %>% 
  pivot_longer(cols = AS:ZN,
               names_to = "Parameter.Symbol",
               values_to = "Value") %>% 
  arrange(Site.ID, Parameter.Symbol, Date)

#using smoothed.yr function
# 5 years
smooth.out <- mapply(Smoothed.Year, 
                     site.para$Site.ID, site.para$Parameter.Symbol, 4.99, 1750)
para.monthly <- data.frame(Site.ID=unlist(smooth.out[1,]),
                           Parameter.Symbol=unlist(smooth.out[2,]),
                           Date=unlist(smooth.out[3,]), 
                           para.yr=unlist(smooth.out[4,]))

dt.pmc.5yr <- para.monthly %>% 
  mutate(Year = floor(Date -0.01),
         Month = round((Date - Year)*12)) %>% 
  select(-Date) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(para.yr)) %>% 
  arrange(Site.ID, Year, Month)

save(dt.pmc.5yr, file = here('data', 'processed', 'pmc-sub-5yr.RDa'))

# 4 years
smooth.out <- mapply(Smoothed.Year, 
                     site.para$Site.ID, site.para$Parameter.Symbol, 3.99, 1400)
para.monthly <- data.frame(Site.ID=unlist(smooth.out[1,]),
                           Parameter.Symbol=unlist(smooth.out[2,]),
                           Date=unlist(smooth.out[3,]), 
                           para.yr=unlist(smooth.out[4,]))

dt.pmc.4yr <- para.monthly %>% 
  mutate(Year = floor(Date -0.01),
         Month = round((Date - Year)*12)) %>% 
  select(-Date) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(para.yr)) %>% 
  arrange(Site.ID, Year, Month)

save(dt.pmc.4yr, file = here('data', 'processed', 'pmc-sub-4yr.RDa'))

dt.pmc.4yr.5yr <- dt.pmc.5yr %>% 
  select(Site.ID, Year, Month) %>% 
  inner_join(dt.pmc.4yr)

save(dt.pmc.4yr.5yr, file = here('data', 'processed', 'pmc-sub-4yr-5yr.RDa'))

# 3 years
smooth.out <- mapply(Smoothed.Year, 
                     site.para$Site.ID, site.para$Parameter.Symbol, 2.99, 1050)
para.monthly <- data.frame(Site.ID=unlist(smooth.out[1,]),
                           Parameter.Symbol=unlist(smooth.out[2,]),
                           Date=unlist(smooth.out[3,]), 
                           para.yr=unlist(smooth.out[4,]))

dt.pmc.3yr <- para.monthly %>% 
  mutate(Year = floor(Date -0.01),
         Month = round((Date - Year)*12)) %>% 
  select(-Date) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(para.yr)) %>% 
  arrange(Site.ID, Year, Month)

save(dt.pmc.3yr, file = here('data', 'processed', 'pmc-sub-3yr.RDa'))

dt.pmc.3yr.5yr <- dt.pmc.5yr %>% 
  select(Site.ID, Year, Month) %>% 
  inner_join(dt.pmc.3yr)

save(dt.pmc.3yr.5yr, file = here('data', 'processed', 'pmc-sub-3yr-5yr.RDa'))

# 2 years
smooth.out <- mapply(Smoothed.Year, 
                     site.para$Site.ID, site.para$Parameter.Symbol, 1.99, 700)
para.monthly <- data.frame(Site.ID=unlist(smooth.out[1,]),
                           Parameter.Symbol=unlist(smooth.out[2,]),
                           Date=unlist(smooth.out[3,]), 
                           para.yr=unlist(smooth.out[4,]))

dt.pmc.2yr <- para.monthly %>% 
  mutate(Year = floor(Date -0.01),
         Month = round((Date - Year)*12)) %>% 
  select(-Date) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(para.yr)) %>% 
  arrange(Site.ID, Year, Month)

save(dt.pmc.2yr, file = here('data', 'processed', 'pmc-sub-2yr.RDa'))

dt.pmc.2yr.5yr <- dt.pmc.5yr %>% 
  select(Site.ID, Year, Month) %>% 
  inner_join(dt.pmc.2yr)

save(dt.pmc.2yr.5yr, file = here('data', 'processed', 'pmc-sub-2yr-5yr.RDa'))

# 1 year
smooth.out <- mapply(Smoothed.Year, 
                     site.para$Site.ID, site.para$Parameter.Symbol, 0.99, 350)
para.monthly <- data.frame(Site.ID=unlist(smooth.out[1,]),
                           Parameter.Symbol=unlist(smooth.out[2,]),
                           Date=unlist(smooth.out[3,]), 
                           para.yr=unlist(smooth.out[4,]))

dt.pmc.1yr <- para.monthly %>% 
  mutate(Year = floor(Date -0.01),
         Month = round((Date - Year)*12)) %>% 
  select(-Date) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(para.yr)) %>% 
  arrange(Site.ID, Year, Month)

save(dt.pmc.1yr, file = here('data', 'processed', 'pmc-sub-1yr.RDa'))

dt.pmc.1yr.5yr <- dt.pmc.5yr %>% 
  select(Site.ID, Year, Month) %>% 
  inner_join(dt.pmc.1yr)

save(dt.pmc.1yr.5yr, file = here('data', 'processed', 'pmc-sub-1yr-5yr.RDa'))

dt.pmc.5yr %>% distinct(Site.ID, Year) %>% nrow() #263 -  820
dt.pmc.4yr %>% distinct(Site.ID, Year) %>% nrow() #303 - 1128
dt.pmc.3yr %>% distinct(Site.ID, Year) %>% nrow() #314 - 1442
dt.pmc.2yr %>% distinct(Site.ID, Year) %>% nrow() #317 - 1745
dt.pmc.1yr %>% distinct(Site.ID, Year) %>% nrow() #317 - 2074
```

### 2021-01-25 code/data-processors/cms-data-processor.Rmd
1 to 5yr mvavg modify smooth function

```{r load data}
dirUSSpatial <- 'C:\\Users\\fkazem01\\Box\\Projects\\USA Spatial\\data\\processed\\'
load(paste0(dirUSSpatial,'ses-site-bz12.RDa'))
load(paste0(dirUSSpatial,'ses-site-bz6.RDa'))
load(paste0(dirUSSpatial,'ses-site-bz24.RDa'))
load(paste0(dirUSSpatial,'brfss-site.RDa'))
load(here('data','processed','master-ndi-pmc.RDa'))
load(here('data','processed','master-ndi-pmc-race.RDa'))
load(here('data','processed','pmc-site-ndi.RDa'))
load(here('data','processed','pmc-1yr.RDa'))
load(here('data','processed','fac-1yr.RDa'))
load(here('data','processed','fac-ex-ec-1yr.RDa'))
load(here('data','processed','pmc-residual-1yr.RDa'))
load(here('data','processed','fac-ex-ec-residual-1yr.RDa'))
```

@@

>>>>

```{r load data}
dirUSSpatial <- 'C:\\Users\\fkazem01\\Box\\Projects\\USA Spatial\\data\\processed\\'
load(paste0(dirUSSpatial,'ses-site-bz12.RDa'))
load(paste0(dirUSSpatial,'ses-site-bz6.RDa'))
load(paste0(dirUSSpatial,'ses-site-bz24.RDa'))
load(paste0(dirUSSpatial,'brfss-site.RDa'))
load(here('data','processed','master-ndi-pmc.RDa'))
load(here('data','processed','master-ndi-pmc-race.RDa'))
load(here('data','processed','pmc-site-ndi.RDa'))
load(here('data','processed','pmc-1yr.RDa'))
load(here('data','processed','fac-1yr.RDa'))
load(here('data','processed','fac-ex-ec-1yr.RDa'))
load(here('data','processed','pmc-residual-1yr.RDa'))
load(here('data','processed','fac-ex-ec-residual-1yr.RDa'))
load(here('data','processed','pmc-sub-5yr.RDa'))
load(here('data','processed','pmc-sub-4yr-5yr.RDa'))
load(here('data','processed','pmc-sub-3yr-5yr.RDa'))
load(here('data','processed','pmc-sub-2yr-5yr.RDa'))
load(here('data','processed','pmc-sub-1yr-5yr.RDa'))
```

@@

```{r PMC - 5yr}
dt.pmc <- dt.pmc.5yr %>% 
  rename_at(vars(AS:ZN), ~paste("Yr", .x, sep = "_"))

write.csv(dt.ndi.pmc, file = here('data', 'processed', 'ndi-pmc-5yr.csv'),
          row.names = F)
```

```{r PMC - 4yr}
dt.pmc <- dt.pmc.4yr.5yr %>% 
  rename_at(vars(AS:ZN), ~paste("Yr", .x, sep = "_"))

write.csv(dt.ndi.pmc, file = here('data', 'processed', 'ndi-pmc-4yr-5yr.csv'),
          row.names = F)
```

```{r PMC - 3yr}
dt.pmc <- dt.pmc.3yr.5yr %>% 
  rename_at(vars(AS:ZN), ~paste("Yr", .x, sep = "_"))

write.csv(dt.ndi.pmc, file = here('data', 'processed', 'ndi-pmc-3yr-5yr.csv'),
          row.names = F)
```

```{r PMC - 2yr}
dt.pmc <- dt.pmc.2yr.5yr %>% 
  rename_at(vars(AS:ZN), ~paste("Yr", .x, sep = "_"))

write.csv(dt.ndi.pmc, file = here('data', 'processed', 'ndi-pmc-2yr-5yr.csv'),
          row.names = F)
```

```{r PMC - 1yr}
dt.pmc <- dt.pmc.1yr.5yr %>% 
  rename_at(vars(AS:ZN), ~paste("Yr", .x, sep = "_"))

write.csv(dt.ndi.pmc, file = here('data', 'processed', 'ndi-pmc-1yr-5yr.csv'),
          row.names = F)
```
